---
title: "analysis"
output: html_document
date: "2024-12-07"
editor_options: 
  markdown: 
    wrap: 72
---

```{r}
library(readxl)
library(dplyr)
library(tidyr)
library(lme4)
library(lmerTest)
```

## Import Data

```{r cars}
df <- read_excel('IFT6253LLMAssistantEvaluation.xlsx', skip=1)

colnames(df)[3] <- "Shot_Example"
colnames(df)[6] <- "Pre_Model"

colnames(df)[7] <- "Class_Precision"
colnames(df)[8] <- "Class_Recall"
colnames(df)[9] <- "Class_F1"

colnames(df)[10] <- "Attr_Precision"
colnames(df)[11] <- "Attr_Recall"
colnames(df)[12] <- "Attr_F1"

colnames(df)[13] <- "Rel_Precision"
colnames(df)[14] <- "Rel_Recall"
colnames(df)[15] <- "Rel_F1"

df$Combination <- factor(df$Combination)
df$Shot_Example <- factor(df$Shot_Example)
df$DSL2Gen <- factor(df$DSL2Gen)
df$Divide <- factor(df$Divide)
df$Pre_Model <- factor(df$Pre_Model)

head(df)
```

## Preprocess

```{r}
df_long <- df %>%
  pivot_longer(
    cols = colnames(df)[7:15], # Select relevant columns
    names_to = c("ElementType", ".value"),
    names_sep = "_"
  )

head(df_long)
```

## Box plots

```{r}
library(ggplot2)

ggplot(df_long, aes(x = Shot_Example, y = Precision, fill = Shot_Example)) +
  geom_boxplot() +
  facet_wrap(~ ElementType) + # Facet by ElementType
  labs(title = "Precision by ElementType and Shot_Example",
    x = "Shot_Example",
    y = "Precision") +
  theme_minimal() +
  theme(legend.position = "none")

ggplot(df_long, aes(x = Shot_Example, y = Recall, fill = Shot_Example)) +
  geom_boxplot() +
  facet_wrap(~ ElementType) + # Facet by ElementType
  labs(title = "Recall by ElementType and Shot_Example",
    x = "Shot_Example",
    y = "Recall") +
  theme_minimal() +
  theme(legend.position = "none")

ggplot(df_long, aes(x = Shot_Example, y = F1, fill = Shot_Example)) +
  geom_boxplot() +
  facet_wrap(~ ElementType) + # Facet by ElementType
  labs(title = "F1 by ElementType and Shot_Example",
    x = "Shot_Example",
    y = "F1") +
  theme_minimal() +
  theme(legend.position = "none")

```

```{r}
ggplot(df_long, aes(x = DSL2Gen, y = Precision, fill = DSL2Gen)) +
  geom_boxplot() +
  facet_wrap(~ ElementType) + # Facet by ElementType
  labs(title = "Precision by ElementType and DSL2Gen",
    x = "DSL2Gen",
    y = "Precision") +
  theme_minimal() +
  theme(legend.position = "none")

ggplot(df_long, aes(x = DSL2Gen, y = Recall, fill = DSL2Gen)) +
  geom_boxplot() +
  facet_wrap(~ ElementType) + # Facet by ElementType
  labs(title = "Recall by ElementType and DSL2Gen",
    x = "DSL2Gen",
    y = "Recall") +
  theme_minimal() +
  theme(legend.position = "none")

ggplot(df_long, aes(x = DSL2Gen, y = F1, fill = DSL2Gen)) +
  geom_boxplot() +
  facet_wrap(~ ElementType) + # Facet by ElementType
  labs(title = "F1 by ElementType and DSL2Gen",
    x = "DSL2Gen",
    y = "F1") +
  theme_minimal() +
  theme(legend.position = "none")
```

```{r}
ggplot(df_long, aes(x = Divide, y = Precision, fill = Divide)) +
  geom_boxplot() +
  facet_wrap(~ ElementType) + # Facet by ElementType
  labs(title = "Precision by ElementType and Divide",
    x = "Divide",
    y = "Precision") +
  theme_minimal() +
  theme(legend.position = "none")

ggplot(df_long, aes(x = Divide, y = Recall, fill = Divide)) +
  geom_boxplot() +
  facet_wrap(~ ElementType) + # Facet by ElementType
  labs(title = "Recall by ElementType and Divide",
    x = "Divide",
    y = "Recall") +
  theme_minimal() +
  theme(legend.position = "none")

ggplot(df_long, aes(x = Divide, y = F1, fill = Divide)) +
  geom_boxplot() +
  facet_wrap(~ ElementType) + # Facet by ElementType
  labs(title = "F1 by ElementType and Divide",
    x = "Divide",
    y = "F1") +
  theme_minimal() +
  theme(legend.position = "none")
```

```{r}
ggplot(df_long, aes(x = Pre_Model, y = Precision, fill = Pre_Model)) +
  geom_boxplot() +
  facet_wrap(~ ElementType) + # Facet by ElementType
  labs(title = "Precision by ElementType and Pre_Model",
    x = "Pre_Model",
    y = "Precision") +
  theme_minimal() +
  theme(legend.position = "none")

ggplot(df_long, aes(x = Pre_Model, y = Recall, fill = Pre_Model)) +
  geom_boxplot() +
  facet_wrap(~ ElementType) + # Facet by ElementType
  labs(title = "Recall by ElementType and Pre_Model",
    x = "Pre_Model",
    y = "Recall") +
  theme_minimal() +
  theme(legend.position = "none")

ggplot(df_long, aes(x = Pre_Model, y = F1, fill = Pre_Model)) +
  geom_boxplot() +
  facet_wrap(~ ElementType) + # Facet by ElementType
  labs(title = "F1 by ElementType and Pre_Model",
    x = "Pre_Model",
    y = "F1") +
  theme_minimal() +
  theme(legend.position = "none")
```

## Check normality for metrics

### Classes

```{r}
shapiro_test_result <- shapiro.test(df_long[df_long$ElementType == "Class", ]$Precision)
print("Precision:")
shapiro_test_result
print("--------------------------------")

shapiro_test_result <- shapiro.test(df_long[df_long$ElementType == "Class", ]$Recall)
print("Recall:")
shapiro_test_result
print("--------------------------------")

shapiro_test_result <- shapiro.test(df_long[df_long$ElementType == "Class", ]$F1)
print("F1:")
shapiro_test_result
print("--------------------------------")
```

Non normal: - Precision

### Attr

```{r}
shapiro_test_result <- shapiro.test(df_long[df_long$ElementType == "Attr", ]$Precision)
print("Precision:")
shapiro_test_result
print("--------------------------------")

shapiro_test_result <- shapiro.test(df_long[df_long$ElementType == "Attr", ]$Recall)
print("Recall:")
shapiro_test_result
print("--------------------------------")

shapiro_test_result <- shapiro.test(df_long[df_long$ElementType == "Attr", ]$F1)
print("F1:")
shapiro_test_result
print("--------------------------------")
```

Non normal: - none

### Rel

```{r}
shapiro_test_result <- shapiro.test(df_long[df_long$ElementType == "Rel", ]$Precision)
print("Precision:")
shapiro_test_result
print("--------------------------------")

shapiro_test_result <- shapiro.test(df_long[df_long$ElementType == "Rel", ]$Recall)
print("Recall:")
shapiro_test_result
print("--------------------------------")

shapiro_test_result <- shapiro.test(df_long[df_long$ElementType == "Rel", ]$F1)
print("F1:")
shapiro_test_result
print("--------------------------------")
```

Non normal: - none \## Analysis \### Class \#### Precision

```{r}
library(lme4)
library(lmerTest)
library(car)
library(emmeans)

df <- df_long[df_long$ElementType == "Class", ]
head(df)

class_precision <- lm(Precision ~ Divide * Pre_Model + DSL2Gen + Shot_Example, data = df)
summary(class_precision)
print("----------------------")

qqnorm(residuals(class_precision))
qqline(residuals(class_precision),col="red")

leveneTest(residuals(class_precision) ~ Divide * Pre_Model, data=df)
print("----------------------")

print("Pre_Model")
emmeans(class_precision, pairwise ~ Pre_Model)$contrasts
print("Divide")
emmeans(class_precision, pairwise ~ Divide)$contrasts
print("Pre_Model | Divide")
emmeans(class_precision, pairwise ~ Pre_Model | Divide)$contrasts
```

#### Recall

```{r}
df <- df_long[df_long$ElementType == "Class", ]
head(df)

class_recall <- aov(Recall ~ Pre_Model * Divide + DSL2Gen + Shot_Example, data = df)
anova(class_recall)
print("----------------------")

print("Pre_Model")
emmeans(class_recall, pairwise ~ Pre_Model)$contrasts
print("Divide")
emmeans(class_recall, pairwise ~ Divide)$contrasts
print("Pre_Model | Divide")
emmeans(class_recall, pairwise ~ Pre_Model | Divide)$contrasts
```

#### F1

```{r}
df <- df_long[df_long$ElementType == "Class", ]
head(df)

class_F1 <- aov(F1 ~ Divide * Pre_Model + DSL2Gen + Shot_Example, data = df)
summary(class_F1)
print("----------------------")

print("Pre_Model")
emmeans(class_F1, pairwise ~ Pre_Model)$contrasts
print("Divide")
emmeans(class_F1, pairwise ~ Divide)$contrasts
print("Pre_Model | Divide")
emmeans(class_F1, pairwise ~ Pre_Model | Divide)$contrasts
```

#### Significant changes

-   Precision
    -   Pre_Model: NOTE: Results may be misleading due to involvement in
        interactions
        -   Pre_Model0 - Pre_Model1 **0.0823 / 0.0322**
    -   Divide: NOTE: Results may be misleading due to involvement in
        interactions
        -   Auto - Manual 0.0502 / 0.4877
        -   Auto - None -0.0432 / 0.5837
        -   Manual - None -0.0934 / 0.1062
    -   Pre_Model \| Divide:
        -   Divide = Auto:
            -   Pre_Model0 - Pre_Model1 **0.1899 / 0.0065**
        -   Divide = Manual:
            -   Pre_Model0 - Pre_Model1 0.0569 / 0.3623
        -   Divide = None:
            -   Pre_Model0 - Pre_Model1 0.0000 / 1.0000
-   Recall
    -   Pre_Model: NOTE: Results may be misleading due to involvement in
        interactions
        -   Pre_Model0 - Pre_Model1 -0.000694 / 0.9845
    -   Divide: NOTE: Results may be misleading due to involvement in
        interactions
        -   Auto - Manual 0.03438 / 0.7100
        -   Auto - None 0.00677 / 0.9865
        -   Manual - None -0.02760 / 0.8004
    -   Pre_Model \| Divide:
        -   Divide = Auto:
            -   Pre_Model0 - Pre_Model1 0.0552 / 0.3787
        -   Divide = Manual:
            -   Pre_Model0 - Pre_Model1 -0.0573 / 0.3614
        -   Divide = None:
            -   Pre_Model0 - Pre_Model1 0.0000 / 1.0000
-   F1
    -   Pre_Model: NOTE: Results may be misleading due to involvement in
        interactions
        -   Pre_Model0 - Pre_Model1 0.0322 / 0.3014
    -   Divide: NOTE: Results may be misleading due to involvement in
        interactions
        -   Auto - Manual 0.05176 / 0.3638
        -   Auto - None 0.00485 / 0.9905
        -   Manual - None -0.04690 / 0.4317
    -   Pre_Model \| Divide:
        -   Divide = Auto:
            -   Pre_Model0 - Pre_Model1 0.110 / 0.0521
        -   Divide = Manual:
            -   Pre_Model0 - Pre_Model1 -0.013 / 0.8071
        -   Divide = None:
            -   Pre_Model0 - Pre_Model1 0.0000 / 1.0000

### Attr

#### Precision

```{r}
df <- df_long[df_long$ElementType == "Attr", ]
head(df)

attr_precision <- aov(Precision ~ Divide * Pre_Model + DSL2Gen + Shot_Example, data = df)
summary(attr_precision)
print("----------------------")

print("Pre_Model")
emmeans(attr_precision, pairwise ~ Pre_Model)$contrasts
print("Divide")
emmeans(attr_precision, pairwise ~ Divide)$contrasts
print("Pre_Model | Divide")
emmeans(attr_precision, pairwise ~ Pre_Model | Divide)$contrasts
```

#### Recall

```{r}
library(emmeans)

df <- df_long[df_long$ElementType == "Attr", ]
head(df)

attr_recall <- aov(Recall ~ Divide * Pre_Model + DSL2Gen + Shot_Example, data = df)
summary(attr_recall)
print("----------------------")

print("Pre_Model")
emmeans(attr_recall, pairwise ~ Pre_Model)$contrasts
print("Divide")
emmeans(attr_recall, pairwise ~ Divide)$contrasts
print("Pre_Model | Divide")
emmeans(attr_recall, pairwise ~ Pre_Model | Divide)$contrasts
```

#### F1

```{r}
library(emmeans)

df <- df_long[df_long$ElementType == "Attr", ]
head(df)

attr_F1 <- aov(F1 ~ Divide * Pre_Model + DSL2Gen + Shot_Example, data = df)
summary(attr_F1)
print("----------------------")

print("Pre_Model")
emmeans(attr_F1, pairwise ~ Pre_Model)$contrasts
print("Divide")
emmeans(attr_F1, pairwise ~ Divide)$contrasts
print("Pre_Model | Divide")
emmeans(attr_F1, pairwise ~ Pre_Model | Divide)$contrasts
```

#### Significant changes

-   Precision
    -   Pre_Model: NOTE: Results may be misleading due to involvement in
        interactions
        -   Pre_Model0 - Pre_Model1 0.0337 / 0.5116
    -   Divide: NOTE: Results may be misleading due to involvement in
        interactions
        -   Auto - Manual -0.0911 / 0.3248
        -   Auto - None **-0.2661 / 0.0014**
        -   Manual - None **-0.1750 / 0.0296**
    -   Pre_Model \| Divide:
        -   Divide = Auto:
            -   Pre_Model0 - Pre_Model1 0.0205 / 0.8165
        -   Divide = Manual:
            -   Pre_Model0 - Pre_Model1 0.0445 / 0.6158
        -   Divide = None:
            -   Pre_Model0 - Pre_Model1 0.0360 / 0.6837
-   Recall
    -   Pre_Model: NOTE: Results may be misleading due to involvement in
        interactions
        -   Pre_Model0 - Pre_Model1 **-0.0858 / 0.0293**
    -   Divide: NOTE: Results may be misleading due to involvement in
        interactions
        -   Auto - Manual **0.1198 / 0.0375**
        -   Auto - None 0.0357 / 0.7003
        -   Manual - None -0.0841 / 0.1667
    -   Pre_Model \| Divide:
        -   Divide = Auto:
            -   Pre_Model0 - Pre_Model1 -0.1131 / 0.0873
        -   Divide = Manual:
            -   Pre_Model0 - Pre_Model1 -0.0789 / 0.2222
        -   Divide = None:
            -   Pre_Model0 - Pre_Model1 -0.0655 / 0.3073
-   F1
    -   Pre_Model: NOTE: Results may be misleading due to involvement in
        interactions
        -   Pre_Model0 - Pre_Model1 -0.0227 / 0.5648
    -   Divide: NOTE: Results may be misleading due to involvement in
        interactions
        -   Auto - Manual 0.0146 / 0.9491
        -   Auto - None -0.1204 / 0.0531
        -   Manual - None **-0.1350 / 0.0291**
    -   Pre_Model \| Divide:
        -   Divide = Auto:
            -   Pre_Model0 - Pre_Model1 -0.0292 / 0.6681
        -   Divide = Manual:
            -   Pre_Model0 - Pre_Model1 -0.0192 / 0.7770
        -   Divide = None:
            -   Pre_Model0 - Pre_Model1 -0.0196 / 0.7729

### Rel

#### Precision

```{r}
df <- df_long[df_long$ElementType == "Rel", ]
head(df)

rel_precision <- aov(Precision ~ Divide * Pre_Model + DSL2Gen + Shot_Example, data = df)
summary(rel_precision)
print("----------------------")

print("Pre_Model")
emmeans(rel_precision, pairwise ~ Pre_Model)$contrasts
print("Divide")
emmeans(rel_precision, pairwise ~ Divide)$contrasts
print("Pre_Model | Divide")
emmeans(rel_precision, pairwise ~ Pre_Model | Divide)$contrasts
```

#### Recall

```{r}
df <- df_long[df_long$ElementType == "Rel", ]
head(df)

rel_recall <- aov(Recall ~ Divide * Pre_Model + DSL2Gen + Shot_Example, data = df)
summary(rel_recall)
print("----------------------")

print("Pre_Model")
emmeans(rel_recall, pairwise ~ Pre_Model)$contrasts
print("Divide")
emmeans(rel_recall, pairwise ~ Divide)$contrasts
print("Pre_Model | Divide")
emmeans(rel_recall, pairwise ~ Pre_Model | Divide)$contrasts
```

#### F1

```{r}
df <- df_long[df_long$ElementType == "Rel", ]
head(df)

rel_f1 <- aov(F1 ~ Divide * Pre_Model + DSL2Gen + Shot_Example, data = df)
summary(rel_f1)
print("----------------------")

print("Pre_Model")
emmeans(rel_f1, pairwise ~ Pre_Model)$contrasts
print("Divide")
emmeans(rel_f1, pairwise ~ Divide)$contrasts
print("Pre_Model | Divide")
emmeans(rel_f1, pairwise ~ Pre_Model | Divide)$contrasts
```

#### Significant changes

-   Precision
    -   Pre_Model: NOTE: Results may be misleading due to involvement in
        interactions
        -   Pre_Model0 - Pre_Model1 0.0725 / 0.2792
    -   Divide: NOTE: Results may be misleading due to involvement in
        interactions
        -   Auto - Manual 0.1932 / 0.0656
        -   Auto - None -0.0263 / 0.9413
        -   Manual - None **-0.2195 / 0.0346**
    -   Pre_Model \| Divide:
        -   Divide = Auto:
            -   Pre_Model0 - Pre_Model1 0.19804 / 0.0964
        -   Divide = Manual:
            -   Pre_Model0 - Pre_Model1 0.00445 / 0.9688
        -   Divide = None:
            -   Pre_Model0 - Pre_Model1 0.01500 / 0.8952
-   Recall
    -   Pre_Model: NOTE: Results may be misleading due to involvement in
        interactions
        -   Pre_Model0 - Pre_Model1 0.00833 / 0.8641
    -   Divide: NOTE: Results may be misleading due to involvement in
        interactions
        -   Auto - Manual -0.0715 / 0.4596
        -   Auto - None **-0.1889 / 0.0140**
        -   Manual - None -0.1174 / 0.1444
    -   Pre_Model \| Divide:
        -   Divide = Auto:
            -   Pre_Model0 - Pre_Model1 -0.0181 / 0.8305
        -   Divide = Manual:
            -   Pre_Model0 - Pre_Model1 -0.0333 / 0.6933
        -   Divide = None:
            -   Pre_Model0 - Pre_Model1 0.0764 / 0.3710
-   F1
    -   Pre_Model: NOTE: Results may be misleading due to involvement in
        interactions
        -   Pre_Model0 - Pre_Model1 0.034 / 0.4837
    -   Divide: NOTE: Results may be misleading due to involvement in
        interactions
        -   Auto - Manual 0.0332 / 0.8364
        -   Auto - None -0.1356 / 0.0794
        -   Manual - None **-0.1688 / 0.0263**
    -   Pre_Model \| Divide:
        -   Divide = Auto:
            -   Pre_Model0 - Pre_Model1 0.0453 / 0.5886
        -   Divide = Manual:
            -   Pre_Model0 - Pre_Model1 0.0016 / 0.9847
        -   Divide = None:
            -   Pre_Model0 - Pre_Model1 0.0550 / 0.5122
